Model Based Systems Engineering with a Docs-as-Code Approach for the SeaLion CubeSat Project

### 2023-05-30 manuscript/template.tex

Presented here is the systems engineering approach of the SeaLion CubeSat mission architecture.  This includes the modeling language, tools, and technical approach used to facilitate the configuration management, design, specification, and implementation of the SeaLion mission architecture for the flight software using a model-based approach.  Through, model-based systems engineering (MBSE), the SeaLion mission architecture team was able to create models, as opposed to documents, that serve as the authoritative source of truth for the conduction of system engineer activities \cite{architecting_spacecraft}.  These models were used to conduct activities such as design, specification, analysis, verification and validation of the system.  This was done by applying the NASA handbook on systems engineering \cite{nasa_handbook} to CubeSat mission design in efforts to facilitate a top-down design methodology from mission concept to specification of subsystem components, including flight software architecture \cite{asundi13_cubes}.  The approach presented herein had additional intents to make it was as simple and easy to use as possible.  This was done by a filesystem-based modelling language that adheres to expected patterns in agile software engineering (i.e., elements for stakeholder needs, user stories, data structures, etc.) using a lightweight YAML-based syntax.  This article also serves as an expansion of the conference proceedings (copyright held by Kevin Chiu and Sean Marquez) presented at American Institute of Aeronautics and Astronautics (AIAA) SciTech Forum 2023 \cite{scitech_proceeding}.

\subsection{CubeSat}
The CubeSat, originating from California Polytechnic State University in 1999, are a standardized form of nanosatellites.  Nanosatellites are satellites typically defined with a mass of less than 10 kg.  CubeSats, also known as Cube Satellites, are defined by the standardized and modular architecture of 1-Unit (1U) cube with dimensions of 10 cm X 10 cm X 10 cm with a mass of up to 2 kg \cite{cds_rev14}.  They can be scaled to 2-Units (2U), 3-Units (3U), or higher depending on units added as shown in figure \ref{fig:cubesat_family} with sizes designated by the CubeSat specification \cite{cds_rev14}.  This is done by the additional of standardized cube units to upscale the CubeSat.  The ability to scale by modularity gives a highly standardized structure for ease of expansion to provide versatility in functionality.  Due to their small size, mass, and lack of dedicated launch vehicles, CubeSats are typically launched as secondary payloads in conjunction with other larger satellites, informally known as “piggy-backing”.  This greatly decreases the cost of launching a CubeSat which greatly increases the accessibility of inserting objects into space.  As such, low budget groups such as universities have gravitated towards their use with increasing numbers \cite{swartwout_data}.  However, this presents challenges since many of the developers in universities are completely new to spacecraft development or systems engineering in general.

@@

\subsection{SeaLion Mission}
The SeaLion CubeSat mission is a joint project between Old Dominion University (ODU), the United States Coast Guard Academy (USCGA), and the Air Force Institute of Technology (AFIT).  The end goal is to produce a 3U CubeSat consisting of 3 payloads for on-orbit validation.  ODU provided one payload while the USCGA and AFIT provided the other two payloads.  SeaLion was initially planned to launch as a secondary payload on a Northrop Grumman Antares Rocket from Wallops Flight Facility (WFF) during March of 2023 \cite{sealion_cdr}.  The prototype CubeSat model is shown in figure \ref{fig:cubesat_blowup}.  The intended mission profile was to have an no-orbit time for mere days due to its planned very low earth orbit (VLEO) altitude.  Thus, SeaLion was to complete validation of its payloads before either power is lost in its non-rechargeable batteries or the satellite reenters and burns up in Earth's atmosphere.  The predicted on-orbit time was 10 days.  However, weight considerations on the planned Antares caused SeaLion to be moved to a Q4 launch on a Firefly rocket from Vandenberg Space Force Base into a sun synchronous orbit, of 500 miles altitude.  The content presented in this article is based on the prior mission profile from the launch at WFF into VLEO.

@@

The payloads are as follows. The first payload is provided by the USCGA and AFIT, is the Impedance Probe (IP).  The IP is derived from U.S. Naval Research Laboratory's (NRL's) 'Space PlasmADiagnostic suitE' (SPADE) aboard NASA's International Space Station (ISS) where plasma density and temp are computed with alternating current (AC) impedance measurements using an innovative, first of its kind surface mounted dipole radio frequency antenna \cite{sealion_cdr}.  The IP part is shown in figure \ref{fig:IP}.

\begin{figure}[H]
    \includegraphics{assets/impedence.png}
    \caption{Impedence Probe}
	\label{fig:IP}
    \end{figure}
	\noindent   
\unskip

The second payload, provided by the USCGA and AFIT, is the multispectral (Me-S) 'Pixel Sensor' with a with a 450 nm - 1000 nm spectral range \cite{sealion_cdr}.  Its purpose is to provide SeaLion's in-situ spectral data as a baseline.  This baseline will be used for future missions that may require this spectral data.  The Me-S part is shown in figure \ref{fig:Me-S}.

@@

The third payload, provided by ODU, is the deployable composite structure (DeCS).  This payload is a proof-of-concept deployable mechanism and composite boom that is meant to be a platform host to a number of number of applications \cite{sealion_cdr}.  For example, these applications include solar panels, solar sails, drag sails, sensory sails, and magnetometer booms.  Deployment on SeaLion is meant to validate the deployable mechanism for composite boom in the space environment and to validate boom dynamic during and after deployment in orbit.  The DeCS upon deployment is shown in figure \ref{fig:DeCS}.

\begin{figure}[H]
    \includegraphics[width=10.5 cm]{assets/decs.png}
    \caption{DeCS as the black popout booms}
	\label{fig:DeCS}
    \end{figure}
	\noindent   
\unskip

The mission scope of having three payloads require special care and attention to ensure success of the mission.  However, many of the SeaLion project team members are new to spacecraft development and systems engineering.  Thus, the SeaLion mission architecture team took action to provide a mission architecture for the SeaLion team to better organize and direct the efforts of the team, to guide an implementation for the flight software, and to facilitate interface and assembly documentation.

@@

CubeSats were initially conceived as educational tools for space systems engineering \cite{heidt_new}.  Now, their roles have been expanded to not only just educational tools but for observation, technology demonstrations, and research that were previous monopolized by much larger satellites due to the aforementioned low cost of production and launch of these CubeSat satellites.  As such, there has been increasing popularity for CubeSats as seen by the number of launches in figure \ref{fig:launch_data} since year 2000 \cite{swartwout_data} apart from the notable exceptions in the years 2020 and 2021; the author speculates that this downturn is due to the COVID-19 pandemic.  The CubeSat design specification \cite{cds_rev14} as well as the availability of commercial off-the-shelf (COTS) parts and kits have greatly influenced the rise of popularity.  For example, a basic CubeSat kit from Pumpkin can be purchased with a baseline price of as little as 6250 US dollars \cite{pumpkin_cubesat}.  Even the SeaLion CubeSat utilizes many COTS parts as well.  Thus, it has become highly accessible to low-budget groups such as small companies and university groups.  CubeSats have caused the “democratization” of space due to their low cost which has allowed many groups to fly satellites \cite{cubesat_handbook}.

\begin{figure}[H]
    \includegraphics[width=10.5 cm]{assets/launch_data.png}
    \caption{Nanosatellite launch data provided by M. Swartwout as of July 21, 2021 \cite{swartwout_data}}
	\label{fig:launch_data}
    \end{figure}
	\noindent   
\unskip

University groups especially are a large contributor in the overall number of launches of CubeSats yearly.  As of July 27, 2021, alone, there have been 68 CubeSat launches with 40 of them being from university groups (about 58 percent of launches) in the year of 2021; university groups have consistently maintained plurality on total launches \cite{swartwout_data}.  This showcases directly how many university-based CubeSat projects occurred or potentially may occur if trends continue onwards into the future.  However, this poses challenges to many of these projects given the mainly student composed teams which includes the SeaLion mission at ODU.

\subsection{Ensuring CubeSat Success}
The increased number of CubeSats launched also means a greater number of CubeSats from universities being launched as well.  The motivation of this article is to improve the success rate of CubeSat missions from university groups by providing readily available and easily usable tools for university teams.  To further reinforce the need to improve the success rate, the following data is presented in figure \ref{fig:mission_status} which showcases the total successes and failures of CubeSats from universities for the given time periods \cite{reliving_24}.  The data provided is categorized by six different mission statuses of unknown, launch fail, dead on arrival (DOA), early loss, partial mission, and full mission.

\begin{figure}[H]
    \includegraphics[width=10.5 cm]{assets/mission_status.png}
    \caption{Mission status of CubeSat university-class missions provided by Swartwout}
	\label{fig:mission_status}
    \end{figure}
	\noindent   
\unskip

As seen in the preceding figure \ref{fig:mission_status}, failure rates over time appear to be increasing among CubeSats launched from universities.  However, Swartwout notes that highest number of the failures originate from “regular independent” groups with a failure rate of 65 percent at the time of data gathering in 2017 \cite{reliving_24}.  These “regular independent” groups are groups that have fewer than four missions nor designated as a national center of for spacecraft development by its government.

The issue present is that many of the growing number of university groups producing CubeSats lack the resources, training, experience, or methodology to reliably give assurance to their missions.  Often, the majority of the work is done by untrained university students that are, many times, unfamiliar with the system engineering, design methodologies, testing, etc. that are associated with CubeSat development.  The SeaLion team also faced these issues as well.

To address some of these issues, the SeaLion mission architecture team sought to help simplify the development process by providing readily available and easily learnable system engineering approaches and tools.  These provided system engineering approaches and tools include factors such as planning, documentation, project management, and simplifying the process.  Special attention should be given to systems engineering and information exchange for multidisciplinary teams \cite{aalto}.  To showcase these factor's importance, a survey conducted, by University of Bristol, on how to set up CubeSat projects of forty CubeSat groups emphasized the following relevant lessons learned \cite{howtosetup}:

\begin{itemize}
	\item	Planning: Make efforts to “spend a lot of time in the planning stage”.;
	\item	Documentation / Project Management: Groups should have “good documentation of requirements, work done and work to do”.;
	\item	Simplicity: Simply anything you possibly can to increase confidence in success.
\end{itemize}

@@

The goal of the SeaLion CubeSat mission architecture was to capture the data structures and expected behaviors for the development of the flight software.  It has to be done such that it can unambiguously understood well enough to be implemented, as well as provide full traceability and rationale for architectural elements with minimal configuration management overhead \cite{sealion_mission_architecture}.  Thus, the SeaLion CubeSat mission architecture had to achieve the following:

\begin{itemize}
	\item	Ensure templates only contain formatting data (this includes not storing boilerplate text in templates);
	\item	Ensure models are the authoritative source of truth for all artifact content (e.g., artifact structure, meta-data, boilerplate, commentary, discussion, diagrams, tables, etc.);
	\item 	Models should persist on the local filesystem.;
	\item 	Documents should be in plaintext as to be compatible with modern distributed version control system (e.g., Git) and for ease of use.;
	\item 	Documents should be able to sit alongside code and speak to one another.;
	\item 	Documents should be model-based as to have a separation of concerns between content and formatting as well as be both human and machine-readable for querying and generating views.;
\end{itemize}

@@

Proper adoption of a MBSE approach also includes the selection of modeling language and modeling tool.  Considerations when selecting the modeling language and tool was overhead incurred from training the team, the technical overhead of setting up modelling tools, and future adaptability.  Refer to table \ref{tab:language} for modelling language down selection overview.  In addition, the Sealion CubeSat mission architecture team eventually decided to adopt a docs-as-code approach to further enhance the MBSE approach to achieve the listed goals shown above.

\subsection{Model-Based Systems Engineering}
As noted prior that attention should be given to as planning, documentation, project management, and simplifying the process.  That special emphasize should be given to systems engineering and information exchange \cite{aalto}.

Traditional approaches use documents as their authoritative source of truth for conducting system engineering activities \cite{architecting_spacecraft}.  Information in a traditional systems engineering approach today is mostly captured informally, not authored based on a methodology, configuration managed in silo tools, adhocly and infrequently integrated, not easily traceable to its provenance, not properly configuration managed, not properly changed managed, and not effectively shared with stakeholders \cite{caesar_model_based_approach}.  These documents often do not have a living relationship with other documents or to other corresponding elements; thus, changes to one document require manual changes to other documents \cite{ibm_mbse}.  Document-based approaches can exacerbate problems since it lacks point-to-point communication channels as well as lacking methods to enforce consistency \cite{call_herber_2022}.

In contrast, a MBSE approach supports capturing information in a highly structured modeling language, authored based on a methodology, configuration managed in a common tool, highly integrated, traceable to its provenance, and sharing with stakeholders.  MBSE solves the afformentioned issues with document-based approaches by providing models that have direct lines of communication to enforce that consistency.  Models providing the following key advantages over document-based approaches \cite{ibm_mbse}:

\begin{itemize}
	\item	Information is readily communicated and shared within the project.;
	\item	Changes are easily accommodated.;
	\item	Traceability is automated.
\end{itemize}

As an example of the efficiency gains that MBSE can achieve with a side-by-side comparison between MBSE and non-MBSE approaches.  Using an architecting process, of 4,858 information element transfers, noted that all of these transfers were done manually with non-MBSE approaches; however, 13\% of these transfers with automated with MBSE with the potential of up to 81\% should it be used for trade study and peer review tasks \cite{younse_cameron_bradley_2021}.  The SeaLion mission architecture team seeks to take advantage of these efficiency gains that MBSE can achieve.  Usage of MBSE for space-related projects is not without precedent either.  Projects have been taking advantage of MBSE such as the ExoMars mission, Euclid, Galileo, and nanosatellite programs \cite{mazzini}, \cite{esa}, \cite{nottage_corns_2012}.  CubeSat projects have also been using MBSE and have shown to “hold promise of reducing the burden of system engineering tasks” and can “promote uniformity and consistency across future CubeSat models” \cite{kaslow}.  These attributes are important to reduce workload amongst team members and also to facilitate new team members as they join future projects.  Especially useful for universities since students are not available long term due to events such as graduation.

\subsection{Documents-As-Code Approach}
Documents-as-code (Docs-as-code) refers to a philosophy that team members should be writing documentation with the same tools as code \cite{docs_as_code}.  This allows for documentation to updated seamlessly without additional work with document tools (doctools).  The code tools would include version control (e.g., Git), issues trackers, code tools (e.g., Visual Studio Code), etc.  To do so would mean that the “following the same workflows as development teams, and being integrated in the product team. It enables a culture where writers and developers both feel ownership of documentation, and work together to make it as good as possible” \cite{docs_as_code}.  The SeaLion team taking advantage of the aforementioned philosophy would realize the benefits of utilizing the same principles and practices used to manage software, using modern version control tools (e.g., Git), for the configuration management of mission and flight software architecture documentation, and captured in a model-based approach \cite{docs_as_code}.  In addition, these models can stored and used persistently on a local file system without the use of cloud based services.  Proprietary services are not required to generate documentation, modify documentation, or modify models.  Similar approaches have been seen as Structurizr \cite{structurizr} and F Prime Prime (FPP) \cite{f_prime_prime}.  FPP is based on F Prime which is an open-source software framework developed by NASA's Jet Propulsion Laboratory \cite{f_prime}; it is a model-driven approach for producing flight software code that can be compiled directly onto flight hardware.  While docs-as-code has precedent, the methodologies noted are not easily accesible to those without a sufficient background in programming.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Modelling Language and Methodology}
The languages considered were SysML V1 \cite{sys_ml}, SysML V2 \cite{sys_ml2}, PlantUML \cite{plantuml}, and the Mach 30 modelling language (M30ML) \cite{mach30_git} with the mission architecture team conducting a trade study to determine the most suitable.  The down selection criteria that the SeaLion CubeSat mission architecture team has taken into consideration is shown in table \ref{tab:language}.

@@

M30ML was chosen for its lightweight human and machine-readable textual syntax, file-based model interchange support (for persisting models directly on the local filesystem), ability to generate both textual and graphical views, and relatively minimal overhead with modern doctools \cite{mach30_git}.  The lightweight textual syntax and minimal over head is especially important for for a team that had very minimal experience working with such tools.  Other candidate modelling languages lacked in many regards compared to M30ML in these criteria and thus, M30ML was selected.  SysML v2 had a good number of characteristics that M30ML had, however, the lack of minimal overhead with modern doctools prevented its adoption.

\subsection{Ontological Modeling Language}
M30ML was developed using the Ontological Modeling Language (OML) as its basis.  OML is a language that enables defining systems engineering vocabularies and using them to describe systems \cite{oml_language}.  OML, inspired by Web Ontology Language 2 (OWL2) and the Semantic Web Rule Language (SWRL), is meant to be a more gentler and more disciplined method of aforementioned standard for use in systems engineering \cite{oml_language}.  OWL2 does not conform easily to individual modelling rules without tooling support; thus, OML was created.  OML is a tool to improve the speed of modeling and the quality of models while in a more concise and human-friendly high-level external representation \cite{oml_origin_and_rationale}.  However, more recently M30ML's development has been moved from OML to LinkML since it uses a lighter weight toolchain and YAML-based syntax \cite{linkml}.

@@

M30ML is a language for modeling an architecture with YAML-based modeling.  YAML is especially important as a file structure since it is a highly structured, machine queryable, human-readable, lightweight, and line-oriented markup language.  This makes it ideal for document generation use cases as well as use with version control tools like Git.  The simple line by line structure as shown in figure \ref{fig:example_yaml} exemplifies its simplicity.  Users are easily able to read, interpret, and edit documents using the YAML file format so as long they are taught what each line element is.  Doctools such as asciidoctor and bibtex were made compatible easily with minimal technical overhead which was taken advantage of for the submission to the AIAA SciTech 2023 Forum \cite{scitech_proceeding}.  M30ML also provided modeling elements familiar in agile software development, such as stakeholder needs, user stories, data structures, and with relationship elements for defining traceability between modeling elements \cite{mach30_git}.

\begin{figure}[H]
    \includegraphics[width=12 cm]{assets/user_story.png}
    \caption{Example YAML file}
	\label{fig:example_yaml}
    \end{figure}   
\unskip

@@

\begin{figure}[H]
    \includegraphics[width=10.5 cm]{assets/view_beacon.png}
    \caption{View Beacon YAML structure}
	\label{fig:view_beacon}
    \end{figure}
	\noindent   
\unskip

>>>>

Presented here is the systems engineering approach of the SeaLion CubeSat mission architecture.  This includes the modeling language, tools, and technical approach used to facilitate the configuration management, design, specification, and implementation of the SeaLion mission architecture for the flight software using a model-based approach.  Through, model-based systems engineering (MBSE), the SeaLion mission architecture team was able to create models, as opposed to documents, that serve as the authoritative source of truth for the conduction of system engineer activities \cite{architecting_spacecraft}.  These models were used to conduct activities such as design, specification, analysis, verification, and validation of the system.  This was done by applying the NASA handbook on systems engineering \cite{nasa_handbook} to CubeSat mission design in efforts to facilitate a top-down design methodology from mission concept to specification of subsystem components, including flight software architecture \cite{asundi13_cubes}.  The approach presented herein had additional intents to make it as simple and easy to use as possible.  This was done by a filesystem-based modelling language that adheres to expected patterns in agile software engineering (i.e., elements for stakeholder needs, user stories, data structures, etc.) using a lightweight YAML-based syntax.  This article also serves as an expansion of the conference proceedings (copyright held by Kevin Chiu and Sean Marquez) presented at American Institute of Aeronautics and Astronautics (AIAA) SciTech Forum 2023 \cite{scitech_proceeding}.

\subsection{CubeSat}
The CubeSat, originating from California Polytechnic State University in 1999, is a standardized form of nanosatellites.  Nanosatellites are satellites typically defined with a mass of less than 10 kg.  CubeSats, also known as Cube Satellites, are defined by the standardized and modular architecture of a 1-Unit (1U) cube with dimensions of 10 cm X 10 cm X 10 cm with a mass of up to 2 kg \cite{cds_rev14}.  They can be scaled to 2-Units (2U), 3-Units (3U), or higher depending on units added as shown in figure \ref{fig:cubesat_family} with sizes designated by the CubeSat specification \cite{cds_rev14}.  This is done by the addition of standardized cube units to upscale the CubeSat.  The ability to scale by modularity gives a highly standardized structure for ease of expansion to provide versatility in functionality.  Due to their small size, mass, and lack of dedicated launch vehicles, CubeSats are typically launched as secondary payloads in conjunction with other larger satellites, informally known as “piggy-backing”.  This greatly decreases the cost of launching a CubeSat which increases the accessibility of inserting objects into space.  As such, low budget groups such as universities have gravitated towards CubeSats with increasing numbers \cite{swartwout_data}.  However, this presents challenges since many of the developers in universities are completely new to spacecraft development or systems engineering in general.

@@

\subsection{SeaLion Mission}
The SeaLion CubeSat mission is a joint project between Old Dominion University (ODU), the United States Coast Guard Academy (USCGA), and the Air Force Institute of Technology (AFIT).  The end goal is to produce a 3U CubeSat consisting of 3 payloads for on-orbit validation.  ODU provided one payload while the USCGA and AFIT provided the other two payloads.  SeaLion was initially planned to launch as a secondary payload on a Northrop Grumman Antares Rocket from Wallops Flight Facility (WFF) during March of 2023 \cite{sealion_cdr}.  The prototype CubeSat model is shown in figure \ref{fig:cubesat_blowup}.  The intended mission profile was to have an on-orbit time for mere days due to its planned very low earth orbit (VLEO) altitude.  Thus, SeaLion was to complete validation of its payloads before either power is lost in its non-rechargeable batteries or the satellite reenters and burns up in Earth's atmosphere.  The predicted on-orbit time was 10 days.  However, mass considerations on the planned Antares rocket caused SeaLion to be moved to a Q4 2023 launch on a Firefly rocket from Vandenberg Space Force Base into a sun synchronous orbit, of 500 miles altitude.  The content presented in this article is based on the prior mission profile from the launch at WFF into VLEO.

@@

The first payload, provided by the USCGA and AFIT, is the Impedance Probe (IP).  The IP is derived from U.S. Naval Research Laboratory's (NRL's) 'Space PlasmADiagnostic suitE' (SPADE) aboard NASA's International Space Station (ISS) where plasma density and temperature are computed with alternating current (AC) impedance measurements using an innovative, first of its kind surface mounted dipole radio frequency antenna \cite{sealion_cdr}.  The IP part is shown in figure \ref{fig:IP}.

\begin{figure}[H]
    \includegraphics[width=13.75 cm]{assets/impedence.png}
    \caption{Impedence Probe}
	\label{fig:IP}
    \end{figure}
	\noindent   
\unskip

The second payload, provided by the USCGA and AFIT, is the multispectral (Me-S) 'Pixel Sensor' with a 450 nm - 1000 nm spectral range \cite{sealion_cdr}.  Its purpose is to provide SeaLion's in-situ spectral data as a baseline.  This baseline will be used for future missions that may require this spectral data.  The Me-S part is shown in figure \ref{fig:Me-S}.

@@

The third payload, provided by ODU, is the deployable composite structure (DeCS).  This payload is a proof-of-concept deployable mechanism and composite boom that is meant to be a platform host of a number of applications \cite{sealion_cdr}.  For example, these applications include solar panels, solar sails, drag sails, sensory sails, and magnetometer booms.  Deployment on SeaLion is meant to validate the deployable mechanism for a composite boom in the space environment and to validate boom dynamics during and after deployment in orbit.  The DeCS upon deployment is shown in figure \ref{fig:DeCS}.

\begin{figure}[H]
    \includegraphics[width=13.75 cm]{assets/decs.png}
    \caption{DeCS as the black popout booms}
	\label{fig:DeCS}
    \end{figure}
	\noindent   
\unskip

The mission scope with three payloads requires special care and attention to ensure success of the mission.  However, many of the SeaLion project team members are new to spacecraft development and systems engineering.  In response, experienced team members took action to provide a mission architecture for the SeaLion team to better organize and direct the efforts of the team, to guide an implementation for the flight software, and to facilitate interface and assembly documentation.

@@

CubeSats were initially conceived as educational tools for space systems engineering \cite{heidt_new}.  Now, their roles have been expanded to not only just educational tools but for observation, technology demonstrations, and research that were previous monopolized by much larger satellites due to the low cost of production and launch of these CubeSats.  As such, there has been increasing popularity for CubeSats as seen by the number of launches in figure \ref{fig:launch_data} since the year 2000 \cite{swartwout_data} apart from the notable exceptions in the years 2020 and 2021; the author speculates that this downturn is due to the COVID-19 pandemic.  The CubeSat design specification \cite{cds_rev14} as well as the availability of commercial off-the-shelf (COTS) parts and kits have greatly influenced the rise of popularity.  For example, a basic CubeSat kit from a space systems company such as Pumpkin can be purchased with a baseline price of as little as 6250 US dollars \cite{pumpkin_cubesat}.  The SeaLion CubeSat also utilizes many COTS parts as well.  Thus, CubeSats have become highly accessible to low-budget groups such as small companies and university groups.  CubeSats have caused the “democratization” of space by allowing many groups to fly satellites \cite{cubesat_handbook}.

\begin{figure}[H]
    \includegraphics[width=13.75 cm]{assets/launch_data.png}
    \caption{Nanosatellite launch data provided by M. Swartwout as of July 21, 2021 \cite{swartwout_data}}
	\label{fig:launch_data}
    \end{figure}
	\noindent   
\unskip

University groups especially are a large contributor in the overall number of launches of CubeSats yearly.  As of July 27, 2021, alone, there have been 68 CubeSat launches with 40 of them being from university groups (about 58 percent of launches) in the year of 2021; university groups have consistently maintained plurality on total launches \cite{swartwout_data}.  This showcases directly how many university-based CubeSat projects occurred or potentially may occur if trends continue onwards into the future.  However, this presents its own challenges.

\subsection{Ensuring CubeSat Success}
The motivation of this article is to improve the success rate of CubeSat missions from university groups by providing readily available and usable tools for university teams.  To further reinforce the need to improve the success rate, the following data is presented in figure \ref{fig:mission_status} which showcases the total successes and failures of CubeSats from universities for the given time periods \cite{reliving_24}.  The data provided is categorized by six different mission success statuses of unknown, launch fail, dead on arrival (DOA), early loss, partial mission, and full mission.

\begin{figure}[H]
    \includegraphics[width=10.5 cm]{assets/mission_status.png}
    \caption{Mission status of CubeSat university-class missions provided by Swartwout \cite{reliving_24}}
	\label{fig:mission_status}
    \end{figure}
	\noindent   
\unskip

As seen in the preceding figure \ref{fig:mission_status}, university CubeSat mission failure rates have increased while partial and full mission success rates have decreased in conjunction with the increasing number of missions as seen in figure \ref{fig:launch_data}.  However, Swartwout notes that the highest number of failures originate from “regular independent” groups with a failure rate of 65 percent at the time of data gathering in 2017 \cite{reliving_24}.  These “regular independent” groups have fewer than four missions performed nor are designated as a national center for spacecraft development by its government.

The issue present is that many of the growing number of university groups producing CubeSats lack the resources, training, experience, or methodology to reliably give assurance to their missions.  The majority of the work is often done by untrained students that are unfamiliar with the aspects of CubeSat development (e.g., system engineering, design methods, and testing).  The SeaLion team also faced these issues as well.

To address some of these issues, SeaLion team members sought to simplify the development process by providing readily available and learnable system engineering approaches and tools.  These provided approaches and tools include factors such as planning, documentation, project management, and simplifying the process.  This is important since special attention should be given to systems engineering and information exchange for multidisciplinary teams \cite{aalto}.  To showcase these factor's importance, a survey of forty CubeSat groups on how to set up CubeSat projects, conducted by the University of Bristol, emphasized the following relevant lessons learned \cite{howtosetup}:

\begin{itemize}
	\item	Planning: Make efforts to “spend a lot of time in the planning stage”.;
	\item	Documentation / Project Management: Groups should have “good documentation of requirements, work done and work to do”.;
	\item	Simplicity: Simplify anything you possibly can to increase confidence in success.
\end{itemize}

@@

The goal of the SeaLion CubeSat mission architecture was to capture the data structures and expected behaviors for the development of the flight software.  The data structures and expected behaviors were captured in such a way that can unambiguously understood well enough to be implemented, as well as provide full traceability and rationale for architectural elements with minimal configuration management overhead \cite{sealion_mission_architecture}.  Thus, the SeaLion CubeSat mission architecture had to achieve the following:

\begin{itemize}
	\item	Ensure templates only contain formatting data (this includes not storing boilerplate text in templates);
	\item	Ensure models are the authoritative source of truth for all artifact content (e.g., artifact structure, meta-data, boilerplate, commentary, discussion, diagrams, tables, etc.);
	\item 	Models should persist on the local filesystem.;
	\item 	Documents should be in plaintext as to be compatible with modern distributed version control system (e.g., Git) and for ease of use.;
	\item 	Documents should be able to persist alongside code and communicate to one another.;
	\item 	Documents should be model-based as to have a separation of concerns between content and formatting as well as be both human and machine-readable for querying and generating views.;
\end{itemize}

@@

Proper adoption of a MBSE approach also includes the selection of the modeling language and modeling tool.  Considerations when selecting the modeling language and tool was overhead incurred from training the team, the technical overhead of setting up modelling tools, and future adaptability.  Refer to table \ref{tab:language} for modelling language down selection overview.  In addition, the Sealion CubeSat mission architecture team eventually decided to adopt a docs-as-code approach to further enhance the MBSE approach to achieve the listed goals shown above.

\subsection{Model-Based Systems Engineering}
As noted prior that attention should be given for planning, documentation, project management, and simplifying the process.  Special emphasize should be given to systems engineering and information exchange \cite{aalto}.

Traditional approaches use documents as their authoritative source of truth for conducting system engineering activities \cite{architecting_spacecraft}.  Information in a traditional systems engineering approach today is mostly captured informally.  For example, it incurs disadvantages such as information not being authored based on a methodology, adhocly and infrequently integrated, not properly configuration managed, not properly changed managed, and not effectively shared with stakeholders \cite{caesar_model_based_approach}.  These documents often do not have a living relationship with other documents or to other corresponding elements; thus, changes to one document require manual changes to other documents \cite{ibm_mbse}.  Document-based approaches can exacerbate problems since it lacks point-to-point communication channels as well as lacking methods to enforce consistency \cite{call_herber_2022}.

In contrast, a MBSE approach captures information in a highly structured modeling language, authored based on a methodology, configuration managed in a common tool, highly integrated, traceable to its provenance, and sharing with stakeholders.  Models provide the following key advantages over document-based approaches \cite{ibm_mbse}:

\begin{itemize}
	\item	Information is readily communicated and shared within the project.
	\item	Changes are easily accommodated.
	\item	Traceability is automated.
\end{itemize}

To showcase the benefits, an architecting process of 4,858 information element transfers was performed.  It noted that all of these transfers were done manually with non-MBSE approaches; however, 13\% of these transfers were automated with MBSE with the potential of up to 81\% should it be used for trade study and peer review tasks \cite{younse_cameron_bradley_2021}.  The SeaLion mission architecture team seeks to take advantage of these efficiency gains that MBSE can achieve.  Space projects have been taking advantage of MBSE such as the ExoMars mission, Euclid, Galileo, and nanosatellite programs \cite{mazzini}, \cite{esa}, \cite{nottage_corns_2012}.  CubeSat projects have also been using MBSE and have shown to “hold promise of reducing the burden of system engineering tasks” \cite{kaslow} and can “promote uniformity and consistency across future CubeSat models” \cite{kaslow}.  These attributes are important to reduce workload amongst team members and to facilitate new team members as they join future projects.  Facilitation of new members is especially useful for universities since students are not available long term due to events such as graduation.

\subsection{Documents-As-Code Approach}
Documents-as-code (Docs-as-code) refers to a philosophy that team members should be writing documentation with the same tools as code \cite{docs_as_code}.  This allows for documentation to be updated seamlessly without additional work with document tools (doctools).  The code tools would include version control (e.g., Git), issues trackers, code tools (e.g., Visual Studio Code), etc.  To do so would mean that writers would follow the same workflows as the development team and they would be integrated into the product team.  A stated result would be to enable "a culture where writers and developers both feel ownership of documentation, and work together to make it as good as possible” \cite{docs_as_code}.  The SeaLion team taking advantage of the aforementioned philosophy would realize the benefits of utilizing the same principles and practices used to manage software, using modern version control tools (e.g., Git), for the configuration management of mission and flight software architecture documentation, and captured in a model-based approach \cite{docs_as_code}.  In addition, these models can be stored and used persistently on a local file system without the use of cloud based services.  Proprietary services are not required to generate documentation, modify documentation, or modify models.  Similar approaches have been seen in Structurizr \cite{structurizr} and F Prime Prime (FPP) \cite{f_prime_prime}.  FPP is based on F Prime which is an open-source software framework developed by NASA's Jet Propulsion Laboratory \cite{f_prime}; it is a model-driven approach for producing flight software code that can be compiled directly onto flight hardware.  While docs-as-code has precedent, the methodologies noted are not easily accesible to those without a sufficient background in programming.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Modelling Language and Methodology}
The languages considered were SysML V1 \cite{sys_ml}, SysML V2 \cite{sys_ml2}, PlantUML \cite{plantuml}, and the Mach 30 modelling language (M30ML) \cite{mach30_git} with the mission architecture team conducting a trade study to determine the most suitable one.  The down selection criteria that the SeaLion CubeSat mission architecture team has taken into consideration is shown in table \ref{tab:language}.

@@

M30ML was chosen for its lightweight human and machine-readable textual syntax, file-based model interchange support (for persisting models directly on the local filesystem), ability to generate both textual and graphical views, and relatively minimal overhead with modern doctools \cite{mach30_git}.  The lightweight textual syntax and minimal overhead is especially important for for a team that has very minimal experience working with such tools.  Other candidate modelling languages lacked in many regards compared to M30ML in these criteria and thus, M30ML was selected.  SysML v2 had a good number of characteristics that M30ML had, however, the lack of minimal overhead with modern doctools prevented its adoption.  At the time of publication of this article, the current state of the art MBSE languages (e.g., SysML v2) have not prescribed an approach for file-based model interchange for persisting models on the local filesystem.  SysML v1 has XML Metadata Interchange (XMI) as a file-based interchange but it can only handle graphical views and not textual views.  Adapting other MBSE langauges would take significent work to adopt a docs-as-code approach in their current states.

\subsection{Ontological Modeling Language}
M30ML was developed using the Ontological Modeling Language (OML) as its basis.  OML is a language that enables defining systems engineering vocabularies and using them to describe systems \cite{oml_language}.  OML, inspired by Web Ontology Language 2 (OWL2) and the Semantic Web Rule Language (SWRL), is meant to be a more gentler and more disciplined method of the aforementioned standard for use in systems engineering \cite{oml_language}.  OWL2 does not conform easily to individual modelling rules without tooling support; thus, OML was created.  OML is a tool to improve the speed of modeling and the quality of models while in a more concise and human-friendly high-level external representation \cite{oml_origin_and_rationale}.  However, more recently, M30ML's development has been moved from OML to LinkML since it uses a lighter weight toolchain and YAML-based syntax \cite{linkml}.

@@

M30ML is a language for modeling an architecture with YAML-based modeling.  YAML as a file type is a highly structured, machine queryable, human-readable, lightweight, and line-oriented markup language.  This makes it ideal for document generation use cases as well as use with version control tools like Git.  The simple line by line structure as shown in figure \ref{fig:example_yaml} exemplifies its simplicity.  Users are readily able to read, interpret, and edit documents using the YAML file format so as long they are taught what each line element is.  Doctools such as asciidoctor and bibtex were made compatible with minimal technical overhead which was taken advantage of for the submission to the AIAA SciTech 2023 Forum \cite{scitech_proceeding}.  M30ML also provided modeling elements familiar in agile software development, such as stakeholder needs, user stories, data structures, and with relationship elements for defining traceability between modeling elements \cite{mach30_git}.

\begin{figure}[H]
    \includegraphics[width=13.75 cm]{assets/user_story.png}
    \caption{Example YAML file}
	\label{fig:example_yaml}
    \end{figure}   
\unskip

@@

\begin{figure}[H]
    \includegraphics[width=13.75 cm]{assets/view_beacon.png}
    \caption{View Beacon YAML structure}
	\label{fig:view_beacon}
    \end{figure}
	\noindent   
\unskip
